|

### 2023-06-14 paper/article.qmd
finished methods

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  fig.align = "center",
  fig.path = "plots/"
)
source("../misc/functions.R")
```

@@

@@

In its most advanced, and probably broadest, conceptualisation, @mcleodSocialControlSocial1999 state that the *protest paradigm* combines *unfavourable story framing*, *reliance on official sources and official definitions* and the *invocation of public opinion* (e.g., the protesters are a minority).

<!-- "Protest coverage research needs to move beyond the dichotomous positive versus negative framing to account for the ways in which violence can in fact be a legit- imate way to attract media attention and legitimize protesters’ grievances and tactics (Cammaerts 2013; Drury and Reicher 2018; Saavedra and Drury 2019)" -->

@@

@@

\newpage
# References

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  fig.align = "center",
  fig.path = "plots/"
)
source("../misc/functions.R")
library(tidyverse); theme_set(theme_minimal())
library(kableExtra)
```

@@

To test these hypotheses, I combine two datasets.
The first was created through content analysis and contains measures of framing in reports about a large number of events in UK mainstream newspapers from 1992 to the end of 2017.
The second one contains information about protest events.
These data are matched and the hypotheses are tested through multilevel logistic regression models, with the presence or absense of a frame as dependant, and the above described factors that are assumed to shape protest reporting as independent variables.

@@

I use media coverage of protests as the dependent variable.
Newspaper articles were employed instead of alternatives, such as online news, social media content or TV transcripts for two reasons.
Firstly, despite declining audiences, mainstream news media such as newspapers are still deemed the most important platforms for public debate and the creation of the public agenda [@rogstadTwitterJustRehashing2016].
Newspapers might have lost some of their intermedia agenda-setting power due to being slower to release breaking news, but the more in-depth writing produced by newspapers nevertheless forms a crucial share of what is redistributed via online news and social media and is key in initiating, amplifying and sustaining attention to an issue [@Langerwindrush2020; @harderIntermediaAgendaSetting2017; @rogstadTwitterJustRehashing2016].
Secondly, suitable data from UK newspapers were already archived and digitally available for the whole period of interest.

Table @tab:outlets presents the selection of newspaper outlets, strategically chosen to contain variety on the two important factors: 
ideology and type of newspaper in the UK
Additionally the selection was made to include the outlets with the highest circulation figures.
I retrieved articles from LexisNexis, utilizing keywords such as "protest" and "demonstration" and several several misspellings.
A pilot ensured the comprehensiveness of these keywords by testing a range of synonyms like "march", "rally", or "direct action", which, however, did not yield any additional relevant articles.
The articles underwent a processing phase using the R-package LexisNexisTools [@gruberlnt2021] which consisted of removing duplicate entries, items discussing non-domestic protests, and false positives.
To maintain theoretical integrity, we opted to limit our data to domestic protests, as studies have indicated notable differences in the coverage of non-domestic protests [@boyleAdherenceProtestParadigm2012].
This led to the final dataset, comprising 27,496 newspaper articles.

```{r outlets}
#| label: outlets
outlets <- tibble::tribble(
  ~Type,        ~Ideology,  ~Outlet,                 ~"Available from", ~"Circulation (2005 -- 2017)",
  "Tabloid",    "Left",    "The Mirror",             1995,              "1,720,000 -- 692,295",
  "Tabloid",    "Right",   "The Sun",                2000,              "3,273,000 -- 1,602,320",
  "Tabloid",    "Right",   "Daily Mail",             1992,              "2,426,000 -- 1,442,924",
  "Broadsheet", "Left",    "The Guardian",           1992,              "366,645 -- 153,431",
  "Broadsheet", "Left",    "The Independent",        1992,              "263,595 -- NA",
  "Broadsheet", "Right",   "Financial Times",        1992,              "419,386 -- 190,046",
  "Broadsheet", "Right",   "Daily Telegraph",        1992,              "460,585 -- 907,329",
  "Broadsheet", "Right",   "The Times",              1992,              "679,190 -- 440,736"
)

footnotes = c(
  "Source for circulation numbers: http://www.magforum.com/papers/nationals.htm",
  "No figures available for The Independent in 2017 as the outlet moved to online-only publication in 2016."
)

outlets |>  
  knitr::kable(format = "latex",
        booktabs = TRUE,
        linesep = "",
        caption = "Selection of Newspaper Outlets for Data Collection",
        longtable = FALSE,
        escape = FALSE) |>  
  kableExtra::add_footnote(footnotes, notation = "alphabet", threeparttable = TRUE)
```

This dataset was coded for framing of protest reporting.
Specifically, I re-use the measurement of framing in newspaper reports on protests and demonstrations from @gruberTrouble2023.
This results in one binary variable per frame: which takes the value 1 if a frame is present and 0 otherwise.
One article can contain several frames, or none.
In short, I frames were identified by coding the frame elements found in @entman1993's definition of framing (problem definition, causal interpretation, moral evaluation, and treatment recommendation) in 500 randomly selected artiocle and combining the coded elements to frames using factor analysis.
They were then coded using supervised machine learning models trained on the manually coded material.

<!-- In its most advanced, and probably broadest, conceptualisation, @mcleodSocialControlSocial1999 state that the *protest paradigm* combines *unfavourable story framing*, *reliance on official sources and official definitions* and the *invocation of public opinion* (e.g., the protesters are a minority). -->
<!-- "Protest coverage research needs to move beyond the dichotomous positive versus negative framing to account for the ways in which violence can in fact be a legit- imate way to attract media attention and legitimize protesters’ grievances and tactics (Cammaerts 2013; Drury and Reicher 2018; Saavedra and Drury 2019)" -->

@@

This article tests the relation between framing of protest and eight independent variable identified from the literature above:
*Protest Goal*, *Protester Violence*, *State Repression*, *Newspaper Ideology*, *Newspaper Type*, *Ideological Divide*, *Days Since Start*, and *Year of Protest*.
*Newspaper Ideology* were obtained through the meta information of the coded news articles *Newspaper Type*.
For all other variables the *Mass Mobilization Project*'s [@ClarkMMPdata2019] dataset on protest events was employed.
This dataset was chosen for pratical reasons as it contains the variables neccesary and covers the chosen time frame (1992-2017).
@ClarkMMPdata2019 use the keywords "Protest", "Demonstration", "Riot", and "Mass Mobilization" on the *LexisNexis* newspaper archive, selecting "All News" as the source.
For the selected time frame in the UK, the *Mass Mobilization Project* lists 471 events between 1992 and the end of 2017, as shown in @fig-MMPyear.

```{r}
#| label: fig-MMPyear
#| fig-cap: "Number of Protest Events in the UK per Year per Clark and Regan (2019)"
readRDS("../misc/protest_event_data_uk.Rds")  |> 
  mutate(year = lubridate::floor_date(date_start, "year"))  |> 
  group_by(year)  |> 
  count()  |> 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = jbg_pal(sel = "grey")) +
  scale_x_date(labels = lubridate::year, breaks = "year",
               expand = c(0, 0)) +
  #scale_y_continuous(labels = scales::percent, limits = c(0, 0.7)) +
  labs(x = NULL, y = NULL, colour = NULL) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.minor = element_blank())
```

To merge the two datasets, I employed a two-staged strategy:
first, I selected all articles in the newspaper dataset that were published no earlier than seven days before an event started and no later than 30 days after it ended;
then I used the variables "Protest Locations", "Protest Group Identity" and "Protester Demands" to extract suitable keywords [see Appendix @app-eventdata].
I then used keywords to filter the articles selected in the first stage:
articles were discarded if they did not feature at least one of the keywords.
This selection was manually validated and overlaps between events were separated by hand.

For the variable *ideological divide*, additional manual coding was necessary:
An ideological divide is present when the newspaper ideology is right-leaning and the protest ideology is coded as left --- or the other way around.
The newspaper ideology is known; however, I had to code the ideology of the protest myself based on the information in the MMP dataset and in other articles about an event.^[See Appendix @app-eventdata for the specific coding.]
Note, that since "right"- and "left"-leaning are concepts which are often disputed, I tried to assess whether a protest aligns closer with the main UK right-wing or left-wing party (i.e., closer to Labour or Conservative politics), as this is how British newspapers are usually divided [@kuhnPoliticsMediaBritain2007].

Using the variables listed in @tab-variables, I fitted multilevel logistic regression models (one for each frame) to establish the influence of the different factors on reporting.
Specifically, I used the *lme4* package in R [@lme4].
This is appropriate as the dependent variables for all models are binary and the articles published in the same newspaper and in the same year effectivly consitute panel dataset since observations must be seen as nested and not independent.
It was considered to use newspapers as the second level variable as unobserved and unobservable heterogeneity is most likely present between outlets.
However, the number of newspapers in the dataset (eight) is considered problematically low for fitting a valid multilevel model [e.g., @bryanmultilevel2016].
Therefore, I used newspaper-year as the Level 2 variable ($n = 180$).
<!-- That means that *The Guardian*, for example, is treated as a different second-level case in 1992, 1993, 1994 and so on.
This makes sense as the last chapter has shown that there is a difference in reporting between all newspaper over time. -->

```{r variables}
#| label: variables
variables <- rio::import("../misc/variables.csv")  |> 
  mutate(across(everything(), ~gsub('"', '', .x, fixed = TRUE)))  |> 
  select(-`Levels-long`, -Mean, -Category)

if (isTRUE(knitr:::pandoc_to() == "latex")) {
  variables  |> 
    custom_df_print(caption = "Model Variables",
                    font_size = 9,
                    footnote = "Reference category",
                    latex_options = "scale_down")  |> 
    pack_rows("Dependent", 1, 1)  |> 
    pack_rows("Event-Level", 2, 4)  |> 
    pack_rows("Outlet-Level", 5, 7)  |> 
    pack_rows("Time-Bound", 8, 9)  |> 
    pack_rows("Level 2", 10, 10)  |> 
    column_spec(5, width = "3.7cm")
} else {
  variables  |> 
    custom_df_print(caption = "Model Variables", font_size = 9)
}
```

@@

This study is not without limitations.
Firstly, it is limited in that it only focuses on newspapers and only addresses the state of protest coverage in the UK.
Some of the interroigated factors for protest coverage might be more or less influential in other forms of media and in oither contexts.

Secondly, the employed *Mass Mobilization Project*'s [@ClarkMMPdata2019] dataset on protest events introduces endogeneity concerns to the study as the data was collected from media reports.
As the dependent variable is media reporting patterns, the fact that some of the independent variables are sourced from this reporting might create a feedback loop complicating our ability to separate cause and effect.
This potentially affects the measurment of the variables *Protest Goal*, *Protester Violence* and *State Repression*.
However, this is risk is currently unavoidable as independent data does not exists.
All of the exiting databases that I could have used alternativly suffer from the same problem, while media-independent information about events is scarce and might be equally biased:
@mccarthyElectronicPrintMedia1999, for example, use police records, which arguably suffer from other biases.
To set this in context, however, research using media-independent information about events has shown that publishing outright false information about an event is relatively rare [@mccarthyElectronicPrintMedia1999].
Reports rather omit information or misrepresent what happened during an event through framing [@earlUseNewspaperData2004] --- like emphasising, for example, violence even when violence was only a very small aspect of what happened.

\newpage
# References {.unnumbered}

::: {#refs}
:::

\newpage
# Appendix {.unnumbered}

```{r article, child = "app-eventdata.qmd"}
```
